---
title: "Smart City Project - Final"
Author: Roberto Batista - rbatista7484@floridapoly.edu
output:
  html_notebook: default
---

#### Loading Libraries 
```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(stats)
library(mlogit)
library(DataExplorer)
library(caret)
library(ggfortify)
```

#### Loading the data
```{r, warning=FALSE, message=FALSE}
setwd("~/Desktop/SMARTCITIES/FINAL/tncs_congestion_empirical_dataset")
year2016 <- read_csv("ESTFILE_2016.csv", col_types = "iccfiifdddffddddddddddddddddddd")
```

#### Exploratory Data Analysis (EDA)

```{r}
year2016 %>% dim()
```

Checking data structure
```{r, eval=TRUE}
str(year2016)
```

```{r, eval=TRUE}
summary(year2016)
```

```{r, eval=TRUE}
year2016 %>% 
  head(3)
```
```{r, eval=TRUE}
plot_missing(year2016)
```

Histograms to check quantitative variables.
```{r, eval=TRUE}
year2016 %>% 
  plot_histogram()
```

Bar plots for qualitative variables analysis.

```{r, eval=TRUE}
plot_bar(year2016)
```

```{r, eval=TRUE}
year2016 %>% 
  count(BASE_INRIX_VOL_PRESIDIO) %>% 
  mutate(perc = (n/sum(n)))
```

#### Variables Selection

Let's remove the obvious variables which are not relevant to the model.

Removing variables related to ID's (`X1`, `ID`, `ModifiedTMC`), `YEAR` column, as all the data is related to year 2016, and `BASE_INRIX_VOL_PRESIDIO` which 99.4% of data is `0`.
```{r}
subset_var <- year2016 %>% 
  select(-c(X1, ID, ModifiedTMC, YEAR, BASE_INRIX_VOL_PRESIDIO))
```

Now checking the statistical relevance of the variables.
```{r, eval=TRUE}
test_model <- lm(FF_TIME ~., data = subset_var)
summary(test_model)
```

The test_model summary presents 'NA's for PHF variables, indicating high collinearity with the dependent variable. Let's confirm it checking colinearity using alias().
```{r, eval=TRUE}
alias(test_model)
```

The variable `PHF` has high collinearity and will be removed from the model. In addition, the variables `TNC_PUDO` and `TNC_VOL` will be removed due to be not significant. According to the dictionary the variable `SPEED_20TH` is a devivation from the same data as `INRIX_SPEED` is and will also removed.

```{r}
subset_var <- subset_var %>% 
  select(-c(PHF, TNC_PUDO, TNC_VOL, SPEED_20TH))
#dim(subset_var)
```

```{r, eval=TRUE}
test_model <- lm(FF_TIME ~., data = subset_var)
summary(test_model)
```
Once removed the variables the variable `PRESIDIO` became also statistical not significant and it will be removed.
```{r}
subset_var <- subset_var %>% 
  select(-PRESIDIO)
```

```{r, eval=TRUE}
test_model <- lm(FF_TIME ~., data = subset_var)
summary(test_model)
```

Let's change the `AT` categorical variable factor levels for a better understanding of model_analysis summary.
Check the current levels
```{r}
levels(subset_var$AT)
```

Change the "0.0" level as the first level to change the reference.
```{r}
subset_var$AT <- relevel(subset_var$AT, "0.0")
levels(subset_var$AT)
```

Let's change the `FT2` categorical variable factor levels for a better understanding of model_analysis summary.

Check the current levels
```{r}
levels(subset_var$FT2)
```

Change the "0.0" level as the first level to change the reference.
```{r}
subset_var$FT2 <- relevel(subset_var$FT2, "1")
levels(subset_var$FT2)
```

Computing the LR model with the variables of interest.
```{r, eval=TRUE}
model_analysis <- lm(FF_TIME ~ . , data = subset_var)
summary(model_analysis)
```

```{r, eval=TRUE}
autoplot(model_analysis)
```

```{r, eval=TRUE}
subset_var %>% 
  ggplot(aes(x = CHAMP_LINK_COUNT, y = FF_TIME))+
  geom_boxplot()
```

Removing FF_Time outliers
```{r}
subset_var %>% 
  filter(FF_TIME >1.5) %>% 
  count()
```

```{r}
subset_var <- subset_var %>% 
  filter(FF_TIME <1.5)
```

```{r}
subset_var %>% 
  dim()
```
Removing the outliers .
```{r}
subset_var <- subset_var %>%
  slice(-c(7100, 6969, 2307, 5514, 5914, 7104, 5049, 1310, 3865, 1937, 4772, 4767, 4685,
           4686, 4094, 5123, 4763, 4101, 4764, 5122, 4105, 4663, 4763, 1869, 4715, 4764,
           5122, 4105, 4663, 5116, 4760, 4623, 5114, 4621, 4091, 5110, 4073, 4615, 4297,
           4301, 4750, 5108, 4080, 5105, 4750, 4619, 4089, 4620, 4745, 5105, 4093, 5102,
           4788, 4604, 5098, 5096, 4743, 4277, 4605, 4738, 5096, 4601, 4734, 5092, 4729,
           4600, 5090, 4603, 5090, 4780, 6359, 5890, 5721, 5053, 6354, 5886, 5718, 5716,
           5883, 6350, 5881, 5715, 6347, 5052, 5878, 5713, 6343, 5051, 5875,  571, 6389,
           5050, 5709, 5872, 6385, 6332, 5870, 5708, 6329, 5868, 5707, 6326, 5866, 5706,
           5048, 6322, 5863, 5704, 5861, 5703, 6319, 5047, 5701, 5858, 6315, 6312, 5856,
           5700, 5046, 5853, 5698, 6308, 5697, 5851, 6305, 6302, 5849, 5656, 5045, 6298,
           5846, 5694, 5693, 5844, 6295, 5692, 5842, 6292, 5691, 5840, 6289, 5044, 2304, 
           5782, 5313, 5626, 5766, 4444, 6171, 5685, 4367, 6217, 6222, 5861, 5059, 6219,
           6222, 3228, 5710, 5202, 3715, 5862, 3715, 5012, 5011, 5014, 5012, 4546, 4659,
           5013, 5011, 4510, 5010))
```

```{r}
subset_var <- subset_var %>%
  slice(-c(5000, 5001, 5002, 5003, 5004, 5005, 5006, 5007, 5008, 5009, 5010, 5011, 5012))
```

```{r, eval=TRUE}
model_analysis <- lm(FF_TIME ~ . , data = subset_var)
summary(model_analysis)
```
```{r}
autoplot(model_analysis)
```
```{r}
subset_var %>% 
  dim()
```

With the final version of the dataset let's test prediction models.
---
_______________________________________________

#### COMPUTING LINEAR REGRESSION PREDICTION

Testing Linear Regression Models

for FF_TIME (free flow travel time) Prediction

Separing the data into training and test set.
```{r}
set.seed(2019)
training_samples <- subset_var$FF_TIME %>% 
  createDataPartition(p=0.8, list = FALSE)
train_data <- subset_var[training_samples,  ]
test_data  <- subset_var[-training_samples, ]
```

Computing the model with train_data
```{r}
train_data %>% 
  colnames()
```
______________________________________________
Model 1
```{r}
model_1 <- lm(FF_TIME ~ TOD + CHAMP_LINK_COUNT + ALPHA + BETA + AT + 
                FT2 + LANES + DISTANCE + CAPACITY + FFS  + INRIX_SPEED  + 
                INRIX_TIME + INRIX_VOL + CHAMP_PCE + CHAMP_VOL + AVG_DUR +  
                AVG_DUR_MAJOR_ARTERIALS + AVG_DUR_MINOR_ARTERIALS
                , data = train_data)
summary(model_1)
```
Make predictions with test_data
```{r}
predictions <- model_1 %>% predict(test_data)
```

Model performance

Prediction Error - RMSE
```{r}
RMSE(predictions, test_data$FF_TIME)
```
R-Square
```{r}
R2(predictions, test_data$FF_TIME)
```

```{r}
autoplot(model_1)
```
____________________________________________
Model 19
```{r}
model_19 <- lm(FF_TIME ~ TOD + 
                CHAMP_LINK_COUNT + 
                #ALPHA + 
                #BETA + 
                AT + 
                FT2 + 
                #LANES + 
                DISTANCE + 
                CAPACITY + 
                FFS  + 
                INRIX_SPEED  + 
                INRIX_TIME + 
                INRIX_VOL + 
                CHAMP_PCE + 
                CHAMP_VOL + 
                AVG_DUR +  
                AVG_DUR_MAJOR_ARTERIALS + 
                AVG_DUR_MINOR_ARTERIALS
                , data = train_data)
summary(model_19)
predictions <- model_19 %>% predict(test_data)
```
Make predictions with test_data
```{r}
predictions <- model_19 %>% predict(test_data)
print("Model performance") 
print("Prediction Error - RMSE:")
RMSE(predictions, test_data$FF_TIME)
print("R-Square:") 
R2(predictions, test_data$FF_TIME)
```
```{r}
autoplot(model_19)
```
____________________________________________

Model 20
```{r}
model_20 <- lm(FF_TIME ~ TOD + 
                CHAMP_LINK_COUNT + 
                #ALPHA + 
                #BETA + 
                AT + 
                FT2 + 
                #LANES + 
                DISTANCE + 
                CAPACITY + 
                FFS  + 
                INRIX_SPEED  + 
                INRIX_TIME + 
                INRIX_VOL + 
                CHAMP_PCE + 
                CHAMP_VOL #+ 
                #AVG_DUR +  
                #AVG_DUR_MAJOR_ARTERIALS + 
                #AVG_DUR_MINOR_ARTERIALS
                , data = train_data)
summary(model_20)
predictions <- model_20 %>% predict(test_data)
```
Make predictions with test_data
```{r}
predictions <- model_20 %>% predict(test_data)
print("Model performance") 
print("Prediction Error - RMSE:")
RMSE(predictions, test_data$FF_TIME)
print("R-Square:") 
R2(predictions, test_data$FF_TIME)
```

```{r}
autoplot(model_20)
```
_______________________________________________

Testing the application of log in the predictor variables and the outcome variable, both in train and test datasets.
```{r}
#train_data$FF_TIME    <- log(train_data$FF_TIME)
#test_data$FF_TIME     <- log(test_data$FF_TIME)

#train_data$INRIX_TIME  <- log(train_data$INRIX_TIME)
#test_data$INRIX_TIME  <- log(test_data$INRIX_TIME)

#train_data$INRIX_SPEED <- log(train_data$INRIX_SPEED)
#test_data$INRIX_SPEED <- log(test_data$INRIX_SPEED)

train_data$FFS         <- log(train_data$FFS)
test_data$FFS         <- log(test_data$FFS)
```

```{r}
train_data$FFS %>% 
  plot_histogram()
```
Model 19 Log
```{r}
model_19_log <- lm(log(FF_TIME) ~ TOD + 
                CHAMP_LINK_COUNT + 
                #ALPHA + 
                #BETA + 
                AT + 
                FT2 + 
                #LANES + 
                DISTANCE + 
                CAPACITY + 
                FFS  + 
                INRIX_SPEED  + 
                INRIX_TIME + 
                INRIX_VOL + 
                CHAMP_PCE + 
                CHAMP_VOL + 
                AVG_DUR +  
                AVG_DUR_MAJOR_ARTERIALS + 
                AVG_DUR_MINOR_ARTERIALS
                , data = train_data)
#summary(model_19_log)
predictions <- model_19_log %>% predict(test_data)
```
```{r}
autoplot(model_19_log)
```

Make predictions with test_data
```{r}
predictions <- Model_26_log %>% predict(test_data)
```
Model performance
Prediction Error - RMSE
```{r}
RMSE(predictions, test_data$FF_TIME)
```
R-Square
```{r}
R2(predictions, test_data$FF_TIME)
```